### 2017-06-29 Miscellaneous/About the authors/paper.tex
Added about the authors folder - not in main document

\documentclass[aps,prl,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}
\usepackage{graphbox}
\usepackage{wrapfig}

\begin{document}

\includegraphics[width=\columnwidth]{photo_peter_rohde}

Peter Rohde is an ARC Future Fellow in the Centre for Quantum Software \& Information at the University of Technology Sydney. He obtained his PhD in theoretical physics at the University of Queensland. He has worked at highly-acclaimed international institutes, including the University of Oxford and the Max-Planck Institute. His theoretical proposals have inspired several leading experimental efforts. In his spare time he is a musician, composer, public speaker, adrenaline junkie, and charity worker. He is a very naughty boy and was expelled from preschool.

\includegraphics[width=\columnwidth]{photo_zuen_su}

Zu-En Su is a PhD candidate at the University of Science \& Technology China.

\includegraphics[width=\columnwidth]{photo_heliang_huang}

He-Liang Huang is a PhD candidate at the University of Science \& Technology China.

\includegraphics[width=\columnwidth]{photo_sihui_tan}

Si-Hui Tan is a Research Scientist at the Singapore University of Technology \& Design. She obtained a Bachelor of Science with Honours from Caltech, and subsequently a PhD from MIT. She has received numerous awards, including the MIT Presidential Fellowship.

\includegraphics[width=\columnwidth]{photo_nana_liu}

Nana Liu is a Postdoctoral Research Fellow at the Singapore University of Technology \& Design.

\includegraphics[width=\columnwidth]{photo_ryan_mann}

Ryan Mann is a PhD candidate in the Centre for Quantum Software \& Information at the University of Technology Sydney. He is a better rock-climber than he is physicist, despite being outstanding at the latter.

\includegraphics[width=\columnwidth]{photo_rohit_ramakrishnan}

Rohit Ramakrishnan is a researcher at the Indian Institute of Science.

\includegraphics[width=\columnwidth]{photo_jon_dowling}

Jon Dowling is Professor and Hearne Chair of Theoretical Physics, and Co-Director of the Hearne Institute for Theoretical Physics at Louisiana State University. He has a highly-acclaimed research career, and is author of the popular book \textit{Schr\"odinger's Killer App: Race to Build the World's First Quantum Computer}.

\includegraphics[width=\columnwidth]{photo_bill_munro}

Bill Munro obtained his PhD at the University of Waikato, and has since held numerous prestigious fellowships and appointments. He currently works at NTT Japan, where he is Senior Research Scientist and leader of the theoretical quantum physics research group.

\end{document}