### 2017-09-08 Sections/a_new_frontier_towards_the_singularity.tex
master text from Tim Byrnes - not edited in

%
% A New Frontier - Towards the Singularity
%

\section{A new frontier -- Towards the singularity} \label{sec:outlook} \index{Outlook} \index{A new frontier}

In this section we provide a non-technical outlook on the future quantum internet and its implications, for the benefit of the technically disinterested reader. This section is in the form of several short essays, requiring no technical background knowledge in quantum computation, quantum mechanics, or mathematics.

We acknowledge that while parts of these essays are highly plausible, if not certain, others are highly speculative, but nonetheless based on believable although somewhat futuristic reasoning.

%
% The Era of Quantum Supremacy
%

\input{Sections/essay_the_era_of_quantum_supremacy}

%
% The Economics of the Quantum Internet
%

\input{Sections/essay_the_economics_of_the_quantum_internet}

%
% The Global Virtual Quantum Computer
%

\input{Sections/essay_the_global_virtual_quantum_computer}

%
% Security Implications of the Global Quantum Internet
%

\input{Sections/essay_security_implications_of_the_global_quantum_internet}

%
% Geostrategic Quantum Politics
%

\input{Sections/essay_geostrategic_quantum_politics}

%
% The Quantum Space Race
%

\input{Sections/essay_the_quantum_space_race}

%
% The Quantum Mind
%

\input{Sections/essay_the_quantum_mind}

%
% The Quantum Singularity
%

\input{Sections/essay_the_quantum_singularity}