### 2017-09-21 Tim Byrnes/rmp.tex
Added clock sync. Added more on qkd

\documentclass[aps,rmp,reprint,amsmath,amssymb,graphicx,longbibliography]{revtex4-1}

\usepackage{bm}
\usepackage{graphicx}% Include figure files

\newcommand{\vq}{{\vec q}}
\newcommand{\vf}{v_{\rm F}}
\newcommand{\kf}{k_{\rm F}}
\newcommand{\Phio}{\Phi_0}
\newcommand{\tp}{t_\perp}
\newcommand{\Hca}{\mathcal{H}}
\newcommand{\vk}{{\bf k}}

\begin{document}

\title{The Quantum Internet --- Towards the Singularity}

\maketitle

\bibliography{references}

\end{document}