### 2017-10-03 Junk/Quantum genetic medicine/quantum_genetic_medicine.tex
Added genetic medicine

\documentclass[aps,pra,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}

\newcommand{\comment}[1]{{\color{blue}{\textbf{#1}}}}

\begin{document}

\bibliographystyle{apsrev}

\title{Quantum genetic medicine}

\frenchspacing

\maketitle

%
% Quantum Genetic Medicine
%

\section{Quantum genetic medicine}\index{Genetic medicine}

Until now, medicine\index{Medicine} has largely relied on one-size-fits-all diagnosis and treatment options. Needless to say, everyone's different, manifest in their unique genotype\index{Genotype}. However, with access to individuals' unqiue genetic makeup, the next generation of medicine will become highly personalised, catering for individual genetic differences. People with different genetic predispositions, mutations, or traits will be able to have treatment options tailored to them.\index{Genetic mutations}

With the ability to sequence individuals' genomes extremely cheaply, we will open up entirely new medical possibilities for genetically-personalised treatments -- the era of \textit{genetic medicine}, the next revolution in medicine.

Doing so will require highly complex processing of massive amounts of genetic and drug information, requiring demanding computational resources. We forsee quantum computing as playing a central role in the processing pipeline of genetic drug development, which we present here.

%
% The Human Genome Project
%

\subsection{The human genome project}\index{Genome}\index{Human genome project}

The first complete mapping of the human genome was a major scientific achievement \cite{humanGenomeProject}, opening up entirely new avenues for medical research\index{Medical research} that were previously never possible. It was, however, an extraordinarily expensive undertaking, costing on the order of \$1b.

Since then, genetic sequencing\index{Genetic sequencing} tools have undergone a massive technological transformation and are now available as commodity hardware at price-points accessible to any well-resourced bioscience lab. This now enables sequencing the human genome orders of magnitude cheaper than the first attempt. It is now possible to map the genome for $\sim$\$1,000's.

Following its own technological Moore's Law\index{Moore's Law}, one can reasonably anticipate this process becoming sufficiently cheap that in the near future it will be economically viable for every individual to have their own personal genome fully sequenced and available for medical use.

The major technological transition that has enabled this is the adoption of \textit{short-read} technology\index{Short-reads}. Using this process, a DNA\index{DNA} sequence is not mapped exhaustively from beginning to end, but rather is chemically deconstructed into an enormous number of \textit{short-reads} -- small genetic segments, each on the order of $\sim$50 base-pairs\index{Base-pairs} in length. Having prepared an enormous pool of such short-reads, they are then sequenced in parallel\index{Parallel sequencing}, yielding a database of $N$ short strings, corresponding to small segments of the larger genome. The task then is to reconstruct a complete genome from this data.

To achieve this, there are two approaches that are commonly employed -- \textit{de novo} assembly, and \textit{mapping}.

%
% De Novo Sequencing
%

\subsubsection{\textit{De novo} sequencing}\index{De novo sequencing}

In \textit{de novo} sequencing we treat the short-read data as a jigsaw puzzle that must be reassembled. We define an overlap threshold $n$\index{Overlap threshold}, and upon comparing every string against every other, look at whether their ends overlap consistently by at least $n$ base-pairs. When a match is found, the respective short-reads are merged into a larger \textit{contig}\index{Contig}. This process continues until no further sufficiently-overlapping strings are found. At this point we should, at least in principle, have a fully reassembled genome, or at least very large contigs belonging to it (assuming a sufficiently large pool of short-reads to begin with).

Open-source software for implementing \textit{de novo} assembly of short-read data is available \cite{velvet}\index{Velvet}, with very efficient computational resource requirements. Note that the number of comparisons between short-reads scales only as $O(N^2)$, and employing hash-table\index{Hash-table} representations for short-reads the lookups for each individual comparison can be efficiently implemented in $O(1)$ time.

The \textit{de novo} approach is extremely powerful, as it does not require any genomic reference\index{Genomic reference}. Rather, we can reconstruct a genome \textit{ab initio}, with sufficient read data.

%
% Mapping
%

\subsubsection{Mapping}\index{Mapping}

The failing of \textit{de novo} sequencing is that the pool of short-read data must be sufficiently large that all the pieces in the jigsaw puzzle are present, such that there are no gaps between neighbouring pieces. However, resources are always finite.

The other approach, \textit{mapping}, is to use an existing genome as a reference\index{Reference genome}. Rather than piece short-reads together, we compare them against this reference to infer their relative locations in the genome. When doing so, we allow some error threshold\index{Error threshold} in the matching. Specifically, we require the Hamming distance\index{Hamming distance} between a short-read and an equally-long segment of the reference be below some threshold. The flexiblity offered by this threshold acommodates for mutational differences between the reference and the short-reads\index{Genetic mutations}.

To illustrate the power of this approach, consider the following. Only a miniscule fraction of the genome differs from one individual to the next -- almost all of it is identical. Thus, mapping allows us to identify the mutational differences between individuals. Importantly, unlike \textit{de novo} sequencing, short-reads can be mapped to the reference even with incomplete data. Non-overlapping short-reads can still be successfully mapped, making this approach applicable (but potentially incomplete) even with small datasets.

In the case of cancer diagnosis\index{Cancer diagnosis}, for example, a cancer cell's genetic makeup is virtually identical to that of the individual, modulo some small number of mutations that characterise the cancer. Using the mapping approach we can quickly identify these mutations, hence understanding the genetics of the underlying cancer.

As with \textit{de novo} sequencing, open-source software for efficiently implementing mapping using commodity hardware is available \cite{Bfast etc}.

%
% Genetic Medicine
%

\subsection{Genetic medicine}\index{Genetic medicine}

At its simplest, the design of genetically-tailored medical treatments will require understanding the interactions between drug compounds\index{Drug compounds} and genetic processes at the molecular level. For example, we might desire that a drug modifies gene expression\index{Gene expression} for specific genes, or the transcription\index{Transcription} of DNA to specific proteins\index{Proteins}. In the case of cancer treatment, we may wish to inhibit the function of biochemical processes\index{Biochemical processes} reliant on genes exhibiting particular cancerous mutations, whilst not affecting healthy, unmutated ones.

The drug design process for conventional medicines is an incredibly tedious one, requiring massive latitudinal and longitudinal studies of the effects of drugs on physiological symptoms. However, in the case of individually tailored medicine such studies are clearly not possible. This will require replacing human studies with accurate molecular-level simulations of these chemical processes. At this scale, interactions are inevitably quantum mechanical in nature, which must be accomodated for in simulations.

The key goal is to properly simulate the interaction of candidate drug compounds, ligands, or functional groups with biochemical processes at the molecular level. Classical techniques for such simulations include estimating electron densities using density functional theory\index{Electron density}\index{Density functional theory} (the potential energy surface)\index{Potential energy surface} of compounds, from which useful properties such as bonding affinities\index{Bonding affinity} may be determined.

%
% Quantum Chemistry
%

\subsection{Quantum chemistry}\index{Quantum chemistry}

Despite the existance of classical estimation techniques, molecular-level interactions are necessarily quantum mechanical in nature. For this reason, classical simulation techniques are limited to approximations that ignore realistic and important quantum effects, since this would require exponential classical computational resources.

Using standard quantum chemistry techniques, one can construct accurate Hamiltonians\index{Molecular Hamiltonians} describing the evolution of molecular systems at the quantum level. Efficiently implementing such simulations can then be performed on quantum computers using standard, efficient Hamiltonian simulation quantum algorithms (Sec.~\ref{})\index{Hamiltonian simulation}.

These Hamiltonians can be constructed \textit{ab initio}, combining relevant one- and two-body interactions\index{One-body interactions}\index{Two-body interactions}, such as Coulomb potential\index{Coulomb potential}, kinetic energy\index{Kinetic energy}, spin- and spin-orbit\index{Spin coupling}\index{Spin-orbit coupling} coupling terms appropriately. A net interaction Hamiltonian may be obtained by summing over the different one- ($i$) and two-body (\mbox{$j,k$}) terms for all particles comprising the system,
\begin{align}
	\hat{H}_\mathrm{int} = \sum_i \hat{H}_i + \sum_{j>k} \hat{H}_{j,k}.
\end{align}
Examples for one-body terms include kinetic energy\index{Kinetic energy}, and for two-body terms Coulomb interactions\index{Coulomb interaction}.

%
% Quantum Drug Trial Simulation
%

\subsection{Quantum drug trial simulations}\index{Quantum drug trial simulation}

These approaches borrowed from quantum chemistry allow simulation of the interaction between drug compounds and molecular biochemical interactions. The next step is to perform this process against a huge library of candidate drug compounds\index{Drug compounds} or functional groups\index{Functional groups}, from which, we hope, some candidates will exhibit desired characterisitics, such as bonding affinities\index{Bonding affinity}.

To achieve this, let use construct a classical algorithm for exhaustively constructing and enumerating organic drug molecules or functional groups, up to some cutoff size\index{Cutoff size}. The length of this list grows exponentially with the cutoff size. Next we implement this algorithm unitarily (which is always possible, writing the classical circuit as a reversible one), and construct it as a quantum oracle\index{Oracle} such that every input state (which acts as a pointer\index{Pointers} to an element in the list) yields as output a qubit representation\index{Qubit molecule representation} of the corresponding drug compound.

Having contructed this oracle, we feed the output qubit molecule description into the Hamiltonian simulation algorithm\index{Quantum simulation}, which yields as output a `score' characterising some desired property of the interaction, such as bonding affinity\index{Bonding affinity}. The joint oracle-simulation algorithm is finally embedded within a quantum search subroutine\index{Quantum search algorithm}, which searches over the input space enumerating the set of drug candidates for scores above some desired threshold. This yields a quadratic enhancement in the number of drug candidates that can be simulated in parallel.

%
% In the Cloud
%

\subsection{In the cloud}\index{Cloud quantum computing}

This quantum processing pipeline lends itself well to cloud-based economic models for its implementation. The pipeline comprises several distinct subroutines\index{Subroutines}, which might be outsourced or distributed independently of one another\index{Outsourced quantum computation}\index{Distributed quantum computation}.

An R\&D organisation might specialise in the algorithmic generation of candidate drug compounds, which they would like to retain as a trade secret\index{Trade secrets}, but desire to outsource\index{Outsourced quantum computation} to third-party drug designers\index{Drug designers}, who wish to employ them in their quantum drug trial simulation pipeline\index{Quantum drug trial simulation}. The quantum search\index{Quantum search algorithm} and Hamiltonian simulation\index{Quantum simulation} subroutines might similarly be outsourced or distributed over the cloud to vendors specialising in the implementation of these particular subroutines.

In an era where the genetic composition of every individual is fully characterised, data security will be of utmost importance -- medical confidentiality\index{Medical confidentiality} will be lifted to an entirely new plane when people's genes are at stake. The nefarious uses for obtaining other people's genomes are immense. This provides a perfect example for the value of encrypted quantum computation\index{Encrypted quantum computation} (Sec.~\ref{sec}). If a medical lab is outsourcing some aspects of a computation involving clients' genetics, it is paramount that this be obscured from third-parties performing the computations in the interest of medical confidentiality.

\bibliography{paper}

\end{document}