### 2017-10-05 Sections/quantum_error_correction.tex

%
% Quantum Error Correction
%

\section{Quantum error correction}\index{Quantum error correction (QEC)}

\comment{To do}

As discussed previously in Sec.~\ref{sec:QOS}\index{Quality of service}, quality of service is a major consideration in any quantum network. If we are transmitting quantum states over a quantum channel, there will inevitably be deterioration in the form of decoherence, loss, and other undesireable effects we wish to mitigate. In this section we review some of the essential quantum error correction (QEC) techniques that can be used to achieve this.

%
% Entanglement Purification
%

\subsection{Entanglement purification}\index{Entanglement purification}

\comment{To do}

%
% 3-Qubit Code
%

\subsection{3-qubit code}\index{3-qubit code}

\comment{To do}

%
% 9-Qubit Code
%

\subsection{9-qubit code}\index{9-qubit code}

\comment{To do}

%
% Stabiliser Codes
%

\subsection{Stabiliser codes}\index{Stabiliser codes}

\comment{To do}

%
% Surface Codes
%

\subsection{Surface codes}\index{Surface codes}

\comment{To do}

%
% Topological codes
%

\subsection{Topological codes}\index{Topological codes}

\comment{To do}

%
% Unitary Error Averaging
%

\subsection{Unitary error averaging} \index{Quantum error correction}\index{Unitary error averaging}\label{sec:error_averaging}

\comment{Redundant copies could in principle be distributed}

\comment{To do}

%
% Gate Failure Codes
%

\subsection{Gate failure codes}\index{Gate failure codes}

\comment{To do}

\comment{Cluster states}

%
% Qubit Loss Codes
%

\subsection{Qubit loss codes}\index{Qubit loss codes}

\comment{cite RohdeHaselgrove}

\comment{Located vs unlocated errors}

\comment{Tree horticultural schemere Rudolph et al.}

%
% Continuous-Variable Quantum Error Correction
%

\subsection{Continuous-variable quantum error correction}\index{Continuous-variable quantum error correction}

\comment{To do}