### 2017-10-11 Sections/essay_the_quantum_mind.tex

\subsection{The quantum mind}\index{Quantum mind}

\comment{To do: quantum machine learning and AI. QCs improve training process.}