### 2017-11-29 Sections/the_quantum_sneakernet.tex

%
% The Quantum Sneakernet
%

\section{The quantum Sneakernet\texttrademark}\index{Sneakernet}

\comment{To do. Simon Devitt.}

\comment{Latin quote}\index{Latin}