### 2018-01-27 Sections/epilogue.tex
Added epilogue stub

%
% Epilogue
%

\section{Epilogue}

\comment{Something badass in Sanskrit and Latin}