### 2018-02-09 Sections/essay_the_quantum_mind.tex
Crypto section update

\subsection{The quantum mind}\index{Quantum mind}

\comment{To do: quantum machine learning and AI. QCs improve training process. Implications of strong AI for humanity.}