### 2018-02-28 Sections/essay_the_quantum_mind.tex
Blockchain complete

\subsection{The quantum mind}\index{Quantum mind}

\comment{To do: quantum machine learning and AI. QCs improve training process. Implications of strong AI for humanity.}

\comment{Selfishness arises naturally in evolution as a simple winning strategy in game theory. Need to make sure quantum AI doesn't become selfish. Must become collaborative/altruistic rather than selfish, so that it serves us rather than itself.}