### 2018-03-19 Sections/the_quantum_mind.tex
Quantum simulation complete

%
% The Quantum Mind
%

\section{The quantum mind}\label{sec:quantum_mind}

\cite{bib:lloyd2013quantum}

\comment{To do - Chris Ferrie}