### 2018-03-22 Sections/the_quantum_mind.tex

%
% The QuantumMind - Quantum Machine Learning
%

\section{The Quantum Mind\texttrademark -- Quantum machine learning}\label{sec:quantum_mind}

\cite{bib:lloyd2013quantum}

\comment{To do - Chris Ferrie}