### 2018-04-04 Sections/authors.tex
Formatting, QML outline

%
% Peter P. Rohde
%

\boldauthor{Peter P. Rohde}
\email[To whom correspondence should be addressed: ]{dr.rohde@gmail.com}
\homepage{http://www.peterrohde.org}
\smallaffiliation{Centre for Quantum Software \& Information (QSI), Faculty of Engineering \& Information Technology, University of Technology Sydney, NSW 2007, Australia}
\smallaffiliation{Hearne Institute for Theoretical Physics, Louisiana State University, Baton Rouge, LA 70803, United States}

%
% He-Liang Huang
%

\boldauthor{He-Liang Huang}
\smallaffiliation{Hefei National Laboratory for Physical Sciences at Microscale and Department of Modern Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
\smallaffiliation{CAS Centre for Excellence and Synergetic Innovation Centre in Quantum Information and Quantum Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
\smallaffiliation{CAS-Alibaba Quantum Computing Laboratory, Shanghai 201315, China}
\smallaffiliation{Henan Key Laboratory of Quantum Information \& Cryptography, Zhengzhou Information Science and Technology Institute, Zhengzhou, Henan 450000, China}

%
% Zu-En Su
%

\boldauthor{Zu-En Su}
\smallaffiliation{Hefei National Laboratory for Physical Sciences at Microscale and Department of Modern Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
\smallaffiliation{CAS Centre for Excellence and Synergetic Innovation Centre in Quantum Information and Quantum Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
\smallaffiliation{CAS-Alibaba Quantum Computing Laboratory, Shanghai 201315, China}

%
% Zixin Huang
%

\boldauthor{Zixin Huang}
\smallaffiliation{Department of Physics \& Astronomy, University of Sheffield}

%
% Chris Ferrie
%

\boldauthor{Chris Ferrie}
\smallaffiliation{Centre for Quantum Software \& Information (QSI), Faculty of Engineering \& Information Technology, University of Technology Sydney, NSW 2007, Australia}

%
% Simon Devitt
%

\boldauthor{Simon Devitt}
\smallaffiliation{Centre for Quantum Software \& Information (QSI), Faculty of Engineering \& Information Technology, University of Technology Sydney, NSW 2007, Australia}

%
% Rohit Ramakrishnan
%

\boldauthor{Rohit Ramakrishnan}
\smallaffiliation{Indian Institute of Science, Bangalore, India}

%
% Si-Hui Tan
%

\boldauthor{Si-Hui Tan}
\smallaffiliation{Singapore University of Technology \& Design, 20 Dover Dr, Singapore 138682}

%
% Nana Liu
%

\boldauthor{Nana Liu}
\smallaffiliation{Singapore University of Technology \& Design, 20 Dover Dr, Singapore 138682}

%
% Ryan L. Mann
%

\boldauthor{Ryan L. Mann}
\smallaffiliation{Centre for Quantum Software \& Information (QSI), Faculty of Engineering \& Information Technology, University of Technology Sydney, NSW 2007, Australia}

%
% Samuel I. Marks
%

%\boldauthor{Samuel I. Marks}
%\smallaffiliation{University of Sydney, Australia}

%
% Jonathan P. Dowling
%

\boldauthor{Jonathan P. Dowling}
\smallaffiliation{Hearne Institute for Theoretical Physics and Department of Physics \& Astronomy, Louisiana State University, Baton Rouge, LA 70803, United States}

%
% Tim Byrnes
%

\boldauthor{Tim Byrnes}
\smallaffiliation{State Key Laboratory of Precision Spectroscopy, School of Physical \& Material Sciences, East China Normal University, Shanghai 200062, China}
\smallaffiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
\smallaffiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
\smallaffiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China} 
\smallaffiliation{Department of Physics, New York University, New York, NY 10003, USA}

%
% William J. Munro
%

\boldauthor{William J. Munro}
\smallaffiliation{NTT Basic Research Laboratories, 3-1 Morinosato-Wakamiya, Atsugi, Kanagawa 243-0198, Japan}
\smallaffiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}

%
% Chao-Yang Lu
%

%\boldauthor{Chao-Yang Lu}
%\smallaffiliation{Hefei National Laboratory for Physical Sciences at Microscale and Department of Modern Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
%\smallaffiliation{CAS Centre for Excellence and Synergetic Innovation Centre in Quantum Information and Quantum Physics, University of Science and Technology of China, Hefei, Anhui 230026, China}
%\smallaffiliation{CAS-Alibaba Quantum Computing Laboratory, Shanghai 201315, China}

%
% Christopher Altman
%

%\boldauthor{Christopher Altman}
%\smallaffiliation{???}