### 2018-05-18 Sections/the_quantum_mind.tex
Multiplexed detector fig

%
% The QuantumMind - Quantum Machine Learning
%

\section{The Quantum Mind\texttrademark\,-- Quantum machine learning}\label{sec:quantum_mind}

\cite{bib:lloyd2013quantum}

\comment{To do - Chris Ferrie}

\subsection{Overview}

\subsection{Classical machine learning}

\subsection{Classical machine learning applied to quantum problems}

\subsection{Quantum machine learning}

\subsection{Quantum learning}