### 2018-05-30 Sections/to_the_future_and_beyond.tex
To the future - biology, optimisation beginnings

%
% To The Future & Beyond
%

In this part we highlight some of -- in the minds of the authors -- some of the most exciting and transformative applications for the future quantum internet and its monstrous unified quantum computational power. None of the technologies discussed here have presently been realised on any practical scale, but all have revolutionary potential to transform the global economy, how we interact with the world, and how we live our lives.

However, these predictions on the future of world-changing applications for quantum technology are merely speculation. We believe it is far more likely that the most impactful and reverberating applications for quantum technology likely haven't even been invented yet.

Currently, there is much ongoing research being conducted into new quantum algorithms and protocols, with major developments being made at a rapid pace. Most likely, mankind's quantum future will be far more exciting than our speculations here, much as speculation surrounding early digital computers never anticipated the digital technological revolution and its impact on every aspect of our modern lives.