### 2018-07-29 Sections/a_new_frontier_towards_the_singularity.tex
Added all dropcaps

%
% A New Frontier - Towards the Singularity
%

\section{A new frontier -- Towards the singularity} \label{sec:outlook} \index{Essays} \index{A new frontier}

\dropcap{I}{n} this section we provide a non-technical outlook on the future quantum internet and its implications, for the benefit of the technically disinterested reader, who merely wishes to grasp some of the `big issues'. This section is in the form of several short essays, requiring no technical background knowledge in quantum computation, quantum mechanics, or mathematics.

We acknowledge that while parts of these essays are highly plausible, if not certain, others are highly speculative, but nonetheless based on believable although somewhat futuristic reasoning. At the very least we hope to stimulate the exchange of ideas and their development. We encourage the reader to question the ideas presented in these essays and put forth their own ideas and predictions.

\textit{``There is no harm in doubt and skepticism, for it is through these that new discoveries are made.''} --- Richard Feynman.

%
% The Era of Quantum Supremacy
%

\input{Sections/essay_the_era_of_quantum_supremacy}

\latinquote{Ad astra per alas porci.}

%
% The Global Virtual Quantum Computer
%

\input{Sections/essay_the_global_virtual_quantum_computer}

\latinquote{Kyrie eleison.}

%
% The Economics of the Quantum Internet
%

\input{Sections/essay_the_economics_of_the_quantum_internet}

\latinquote{Deus ex Machina.}

%
% QuantCoin
%

\input{Sections/essay_quantcoin}

\latinquote{\href{https://www.youtube.com/watch?v=lhyaiOZhpSg}{End the Fed!} Libertas justitia veritas. Libertas perfundet omnia luce.}

%
% Security Implications of the Global Quantum Internet
%

\input{Sections/essay_security_implications_of_the_global_quantum_internet}

\latinquote{Annus horribilis.}

%
% Geostrategic Quantum Politics
%

\input{Sections/essay_geostrategic_quantum_politics}

\latinquote{Magnus est mundus.}

%
% The Quantum Space Race
%

\input{Sections/essay_the_quantum_space_race}

\latinquote{Carpe futurum.}

%
% The Quantum Mind
%

\input{Sections/essay_the_quantum_mind}

\latinquote{Benedictus benedicat. Per jesum christum dominum nostrum. Amen.}

%
% The Quantum Singularity
%

%\input{Sections/essay_the_quantum_singularity}

%\latinquote{Non ducor, duco.}