### 2018-08-10 Sections/index.tex
Clock sync, scale free nets

%
% See Only
%

\index{Web-graph \seeonly{Internet web-graph}|gobbleone}
\index{Pareto distribution \seeonly{Power law}|gobbleone}
\index{Secret key \seeonly{Private key}|gobbleone}
\index{Operator-sum representation \seeonly{Kraus operators}|gobbleone}
\index{Trapdoor functions \seeonly{One-way functions}|gobbleone}
\index{Error averaging \seeonly{Unitary error averaging}|gobbleone}
\index{Random graph \seeonly{Percolation graph}|gobbleone}
\index{Entanglement distillation \seeonly{Entanglement purification}|gobbleone}
\index{Holevo bound \seeonly{Holevo quantity}|gobbleone}
\index{Collisions \seeonly{Hash collisions}|gobbleone}
\index{Redundancy code \seeonly{3-qubit code}|gobbleone}
\index{Ligand libraries \seeonly{Drug compound libraries}|gobbleone}
\index{Functional group libraries \seeonly{Drug compound libraries}|gobbleone}
\index{Hamiltonian simulation \seeonly{Quantum simulation}|gobbleone}
\index{Vietoris-Rips simplicial complex \seeonly{Simplicial complex}|gobbleone}
\index{Attack-independent security \seeonly{Information-theoretic security}|gobbleone}
\index{7-qubit code \seeonly{Steane code}|gobbleone}
\index{Bit-flip code \seeonly{3-qubit code}|gobbleone}
\index{Phase-flip code \seeonly{3-qubit code}|gobbleone}
\index{Cryptographic hash functions \seeonly{Hash functions}|gobbleone}
\index{Hashing \seeonly{Hash functions}|gobbleone}
\index{Mining \seeonly{Bitcoin mining}|gobbleone}
\index{Interferometry \seeonly{Optical interferometry, Mach-Zehnder (MZ) interference, Hong-Ou-Mandel (HOM) interference}|gobbleone}
\index{Measurement-based quantum computation \seeonly{Cluster state model}|gobbleone}
\index{Graph states \seeonly{Cluster state model}|gobbleone}
\index{State fidelity \seeonly{Fidelity}|gobbleone}
\index{Process fidelity \seeonly{Fidelity}|gobbleone}
\index{Known unitary errors \seeonly{Unitary errors}|gobbleone}
\index{Unknown unitary errors \seeonly{Unitary errors}|gobbleone}
\index{Relaxation channel \seeonly{Amplitude damping channel}|gobbleone}
\index{Entanglement distillation \seeonly{Entanglement purification}|gobbleone}
\index{Integer factorisation \seeonly{Shor's algorithm}|gobbleone}
\index{Factorisation \seeonly{Shor's algorithm}|gobbleone}
\index{Discrete logarithm algorithm
 \seeonly{Shor's algorithm}|gobbleone}
\index{Period-finding \seeonly{Shor's algorithm}|gobbleone}
\index{Quantum search algorithm\seeonly{Grover's algorithm}|gobbleone}
\index{Tomography \seeonly{Quantum state tomography, quantum process tomography}|gobbleone}
\index{Completely positive maps (CP-maps) \seeonly{Quantum processes}|gobbleone}
\index{Quantum circuits \seeonly{Circuit model}|gobbleone}
\index{Dijkstra's algorithm \seeonly{Shortest-path algorithm}|gobbleone}
\index{A$^*$ algorithm \seeonly{Shortest-path algorithm}|gobbleone}
\index{Dijkstra's algorithm \seeonly{Shortest-path algorithm}|gobbleone}
\index{Switches \seeonly{Optical routers}|gobbleone}
\index{Routers \seeonly{Optical routers}|gobbleone}
\index{Photon wave-packets \seeonly{Wave-packets}|gobbleone}
\index{Frequency filtering \seeonly{Spectral filtering}|gobbleone}
\index{Parity measurements \seeonly{Bell measurements}|gobbleone}
\index{Partial Bell measurements \seeonly{Bell measurements}|gobbleone}
\index{Type-I fusion gates \seeonly{Fusion gates}|gobbleone}
\index{Type-II fusion gates \seeonly{Fusion gates}|gobbleone}
\index{Cluster state fusion strategies \seeonly{Fusion strategies}|gobbleone}
\index{Pauli channels \seeonly{Bit-flip channel, bit-phase-flip channel, dephasing channel, depolarising channel}|gobbleone}
\index{Efficiency \seeonly{Loss channel, detector efficiency, computational efficiency}|gobbleone}
\index{Unstructured database searching \seeonly{Grover's algorithm}|gobbleone}
\index{Time of maturity \seeonly{Maturity}|gobbleone}
\index{Scaling functions \seeonly{Computational scaling functions}|gobbleone}
\index{Qubit tax \seeonly{Taxation}|gobbleone}
\index{Derivatives \seeonly{Forward contract pricing model, forward price}|gobbleone}
\index{Black box \seeonly{Oracle}|gobbleone}
\index{Quantum space race \seeonly{Space race}|gobbleone}
\index{W-function \seeonly{Wigner function}|gobbleone}
\index{Bucket photo-detector \seeonly{Non-number-resolved photo-detector}|gobbleone}
\index{Cryptanalysis \seeonly{Differential cryptanalysis, linear cryptanalysis, frequency analysis}|gobbleone}
\index{Man-in-the-middle attack \seeonly{Intercept-resend attack}|gobbleone}
\index{Linear optics quantum computation (LOQC)\seeonly{Universal linear optics quantum computation}|gobbleone}
\index{Quantum cost vector analysis\seeonly{Cost vector analysis}|gobbleone}
\index{Attenuation\seeonly{Photon loss}|gobbleone}
\index{Quantum genetic medicine\seeonly{Genetic medicine}|gobbleone}
\index{Mutations \seeonly{Genetic mutations}|gobbleone}
\index{Human genome project \seeonly{Genome}|gobbleone}
\index{Potential energy surface \seeonly{Electron density}|gobbleone}
\index{Electrostatic interaction \seeonly{Coulomb interaction}|gobbleone}
\index{Islands \seeonly{Contig islands}|gobbleone}
\index{Time-dependent cost of computation postulate \seeonly{Cost of computation postulate}|gobbleone}
\index{$+$ micro-cluster states \seeonly{Plus micro-cluster states}|gobbleone}
\index{Orlin algorithm \seeonly{Minimum cost flow algorithm}|gobbleone}
\index{Ford-Fulkerson algorithm \seeonly{Maximum flow algorithm}|gobbleone}
\index{Toric code \seeonly{Surface codes}|gobbleone}
\index{Symmetric-key cryptography \seeonly{Private-key cryptography}|gobbleone}
\index{Asymmetric-key cryptography \seeonly{Public-key cryptography}|gobbleone}
\index{Einstein-Podolsky-Rosen (EPR) pairs \seeonly{Bell states}|gobbleone}
\index{Singlet state \seeonly{Bell states}|gobbleone}
\index{Triplet state \seeonly{Bell states}|gobbleone}
\index{Topological codes \seeonly{Surface codes}|gobbleone}

%
% See Also
%

\index{Distributed unitary error averaging @\also{Unitary error averaging}|gobbleone}
\index{Nano-satellite @\also{Micro-satellite}|gobbleone}
\index{Micro-satellite @\also{Nano-satellite}|gobbleone}
\index{Two-channel two-port switches @\also{Optical routers}|gobbleone}
\index{Multiplexers @\also{Optical routers}|gobbleone}
\index{Demultiplexers @\also{Optical routers}|gobbleone}
\index{Linear multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Pyramid multiplexers \& demultiplexers @\also{Optical routers}|gobbleone}
\index{Single-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Multi-channel multi-port switches @\also{Optical routers}|gobbleone}
\index{Quantum error correction (QEC) @\also{3-qubit code, 9-qubit code, topological codes, unitary error averaging}|gobbleone}
\index{Homomorphic encryption @\also{Encrypted quantum computation}|gobbleone}
\index{Blind quantum computation @\also{Encrypted quantum computation}|gobbleone}
\index{Phase-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Polarisation-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Displacement-key encoding @\also{Encrypted quantum computation}|gobbleone}
\index{Non-optical encodings @\also{Matter qubits}|gobbleone}
\index{Toric code @\also{Topological codes}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Distance to horizon @\also{Line-of-sight}|gobbleone}
\index{Line-of-sight @\also{Distance to horizon}|gobbleone}
\index{Micro-cluster states @\also{Star micro-cluster states, plus micro-cluster states, linear micro-cluster states, snowflake micro-cluster states}|gobbleone}
\index{Star micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Plus micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Linear micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Snowflake micro-cluster states @\also{Micro-cluster states}|gobbleone}
\index{Single-qubit quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Time-dependent quantum computational leverage @\also{Quantum computational leverage}|gobbleone}
\index{Quality of service @\also{Quantum error correction, entanglement purification}|gobbleone}

%
% Index
%

\printindex