### 2018-08-15 Junk/QuantCoin/Peter Rohde bibliography/Peter_Rohde_bib.tex

\documentclass[aps,pra,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}

\begin{document}

\bibliographystyle{apsrev}

\title{Peter Rohde Bibliography}

\date{\today}

\frenchspacing

\maketitle

\cite{RohdeCV16, RohdeRecycle16, RohdeLoop15, RohdeLeap15, RohdeLoopMulti15, MORDOR, RohdeWavelet15, RohdeArbSpec15, RohdeUnivLoop15, RohdePhotAdd15, RohdeIntroBS15, RohdeLoop14, RohdeDisp15, RohdeECT14, RohdeAvoid14, RohdeCong16, RohdeCat15, RohdeSPDC13, RohdeQWMem13, RohdeInfCap13, RohdeArbLow12, RohdeIncDim13, RohdeQWExp12, RohdeQWEnc12, RohdeErrBS12, RohdeOptDet11, RohdeEntDynQW12, RohdeTimeRes11, RohdeQWComment10, RohdeMultiWalk11, RohdeRemSpin08, RohdeCSNonLin07, RohdeCompDet07, RohdeImpMulti07, RohdeNoiseEnt07, RohdeCatBreed07, RohdeErrProp07, RohdeStratCS07, RohdeSpecTomo06, RohdeSpecStruc07, RohdeLocUnloc08, RohdeTolTrade07, RohdeModelMM06, RohdeOptEntPur06, RohdeModPhotoDet06, RohdeOptPhot05, RohdeGateChar05, RohdeApproxDet05, RohdeFreqTemp05}

\bibliography{Peter_Rohde_bib}

\end{document}