### 2018-08-22 Sections/acknowledgements.tex
acks

%
% Acknowledgments
%

\section*{Acknowledgments}

\begin{itemize}
\item We acknowledge Gavin Brennen, Joseph Clark, Eric Cavalcanti, Charlotte Bridault, Marco Tomamichel, Jonathan Olson, Samuel Marks \& Darryl Veitch for helpful discussions.
\item We thank the authors of QCircuit (\texttt{\href{http://physics.unm.edu/CQuIC/Qcircuit/}{http://physics.unm.edu/CQuIC/Qcircuit/}}) for providing this very helpful package.
\item P.P.R is funded by an ARC Future Fellowship (project FT160100397).
\item T.B is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (Grant No. 61571301); the Thousand Talents Program for Distinguished Young Scholars (Grant No. D1210036A); and the NSFC Research Fund for International Young Scientists (Grant No. 11650110425); NYU-ECNU Institute of Physics at NYU Shanghai; and the Science \& Technology Commission of Shanghai Municipality (Grant No. 17ZR1443600).
\item J.P.D would like to acknowledge support from the US Air Force Office of Scientific Research, the Army Research Office, the National Science Foundation, and the Northrop-Grumman Corporation.
\item We thank almost unethical quantities of 1,3,7-Trimethylxanthine\index{Caffeine} for aiding in the motivation for this work.
\end{enumerate}