### 2018-10-09 Sections/essay_to_the_future_QKD.tex
Minor tweaks, formatting

\section{The future of quantum cryptography}\label{sec:essay_future_QKD}\index{Quantum key distribution (QKD)}\index{Quantum cryptography}

\dropcap{Q}{uantum} key distribution (QKD) is the first quantum information task to reach commercialisation. Its purpose is to distribute a secret key between two trusted parties who share a quantum channel, as well as a classical channel for authentication. Unlike current cryptography systems, which are secure based on the limitations of an adversary's computer (\textit{computational security}\index{Computational security}), the security of QKD is based on the laws of quantum mechanics, providing guaranteed security unless our understanding of quantum physics is inherently wrong (\textit{information theoretic security}\index{Information theoretic security}).

The typical setting of QKD is as follows. The two trusted parties who want to establish a secret key are typically named Alice and Bob. They share two channels: a quantum channel, which allows them to send quantum state to one another; and, a classical channel, with which they can send classical messages. The communication over the classical channel is assume to be public, and the eavesdropper (Eve) has full anonymous access to it. However, Eve cannot modify messages shared over the classical channel.

The quantum channel is subject to possible manipulation by Eve. The task of Alice and Bob is thus to guarantee security against an adversarial eavesdropper. The typical protocol assumes that Alice and Bob do not share any secret to begin with. The origin of the security of QKD springs from the fundamentals of quantum mechanics, that is, any action by an observer on a quantum state necessarily induces a change in the state. This means that in combination with classical communication, actions of an eavesdropper cannot go undetected.

The ultimate goal of a QKD network is long distance secure quantum communication with imperfect sources.

Despite the significant advances in both the theoretical and experimental development of QKD, a number of challenges remain for QKD to be widely used in securing daily communications \cite{bib:RevModPhys.81.1301, bib:diamanti2016practical}. Experimentally, much effort is being invested into improving the performance of QKD systems. On the theory side,  showing the security of a QKD system with finite key size is also a challenge, because information-theoretic security is achieved only when immunity against the most general (coherent) attack is proven \cite{bib:diamanti2016practical}.

\subsection{Performance}

Some of the criteria for assessing the performance of a QKD scheme include key-rate, distance, cost and robustness.

\subsubsection{Key-rate}

Currently, a strong disparity exists between classical and quantum key distribution rates. Classical optical communication delivers in the order of 100Gbits/second per wavelength, whereas communication rates only on the order of Mbit/s are achieved with QKD.

The obtained key-rate depends on the performance of the detector used. For QKD based on single-photon detection techniques, to achieve a high bit-rate, one requires true single-photon states, in combination with detectors with high efficiency and short dead-time. Current developments are promising, with a reported quantum efficiency of $93\%$ at telecom wavelengths \cite{bib:marsili2013detecting}.

For continuous-variable QKD systems, increasing the bandwidth of the homodyne/heterodyne detectors whilst keeping the electronic noise low is essential.

\subsubsection{Distance}

Extending the range of QKD systems is a major challenge and driving factor for QKD in terms of future network applications. Basically, two approaches are being pursued -- free-space and repeaters. A quantum repeater, similar to its classical analogue, is a device that can extend the range of quantum communication between sender and receiver. However, one cannot amplify the signal that contains the quantum information, owing to the no-cloning theorem\index{No-cloning theorem}, which prohibits making copies of unknown quantum states. The fact that an intercept-resend attack by Eve must disrupt the state of the system is the basis for the security of QKD.

A quantum repeater effectively needs to restore the quantum information without measuring it directly, and is a component which is extremely technologically challenging. Over optical fibre networks, the standard loss for 1550nm is 0.2dB/km. Over long enough distances, this unavoidable loss will reduce the key-rate to a level of little practical relevance, therefore a ground-based solution would be to divide the entire channel into segments, where two partners exchange pairs of entangled photons and store it in a quantum memory \cite{bib:PhysRevLett.81.5932, bib:PhysRevA.59.169}.

The second is to use free-space quantum communication techniques via satellite links. Satellite QKD is achievable with  present-day technology. Here satellites are used as intermediate trusted nodes for communication between locations on the ground. Direct links can be established between ground stations and the satellite, thus enabling communication between parties separated by long distances. Satellite QKD suffers less loss, since the effective thickness of the atmosphere is only around 10km when a satellite is directly overhead. In China, satellite QKD over 1200km has been demonstrated \cite{bib:liao2017satellite}.

\subsubsection{Cost \& robustness}

For QKD systems to be consumer-friendly, low cost and robustness are crucial features. Preferably QKD systems should make use of existing data fibre-optic infrastructure since the use of dark fibres are not only expensive, but often unavailable \cite{bib:diamanti2016practical}. Single-photon detectors at room temperatures are also desirable,  because this can remove the requirement for cryogenic cooling, hence reducing power consumption.

Integrated photonic platforms are being explored to reduce the cost, since miniaturisation can lead to light-weight, low-cost QKD modules that can be mass-manufactured.

Currently, two platforms are being explored: silicon \cite{bib:lim2014review}, and indium phosphide \cite{bib:smit2014introduction}. A reconfigurable QKD system employing an In-P transmitter and silicon detectors has been demonstrated in the laboratory \cite{bib:sibson2017chip}.

\subsection{New protocols}

In parallel to hardware development, research efforts are being directed at finding new QKD protocols which can outperform existing ones. Two of these are high-dimensional (HD) QKD and the Round-Robin differential phase shift protocol (RR-DPS).

High-dimensional QKD aims at encoding more than one bit in each detected photon, which can increase the information capacity when the photon rate is limited. Security proofs against collective attacks are being developed, and an experiment has demonstrated an information capacity 6.9 bits per coincidence rate at 2.7Mbit/s over 20km \cite{bib:zhong2015photon}.

The Round-Robin DPS protocol \cite{bib:sasaki2014practical} removes the need to monitor signal disturbance. In a conventional QKD protocol, the noise parameter needs to be estimated; and if high precision is required, the portion of the signal that is sacrificed increases, thus decreasing the efficiency of the protocol \cite{bib:cai2009finite, bib:hayashi2014security}. This protocol has a high tolerance to the quantum-bit error rate ($< 50\%$) \cite{bib:xu2015discrete}, and makes it attractive for implementation when high systematic errors are unavoidable.

However, currently, neither of the protocols out-compete the more mature decoy-state BB84.

\subsection{Challenge in security}

Although QKD protocols are provably information-theoretically secure, physical implementations often contain imperfections which are not considered in the theoretical model. Attacks can be designed to exploit such imperfections, on either the source or the detectors.

Table.~\ref{tab:attacks} (from \cite{bib:lo2014secure}) summarises some attacks against certain commercial and research systems.

\startnormtable
\begin{table*}[!htbp]
\begin{tabular}{|c|c|c|c|} 
 \hline
 Attack &  Targeted component & Tested system & References\\ 
  \hline
  \hline
Time shift
        & Detector & Commercial & \cite{bib:qi2005time, bib:PhysRevA.78.042333, bib:PhysRevA.74.022313}\\
Time information & Detector & Research & \cite{bib:lamas2007breaking} \\
Detector control & Detector  &   Commercial & \cite{bib:lydersen2010hacking, bib:yuan2010avoiding}\\
Detector control  & Detector  & Research & \cite{bib:gerhardt2011full} \\
Detector dead-time      & Detector  & Research   & \cite{bib:weier2011quantum}      \\
Channel calibration    & Detector  &  Commercial  & \cite{bib:jain2011device}      \\
Phase remapping  &  Phase modulator & Commercial & \cite{bib:xu2010experimental} \\
Phase information & Source & Research & \cite{bib:tang2013source}          \\
Device calibration  & Local oscillator & Research & \cite{bib:jouguet2013preventing} \\
                \hline
\end{tabular}
\caption{\label{tab:attacks} Summary of various attacks against some commercial and 
research QKD systems}
\end{table*}
\startalgtable

To regain security, a number of solutions have been proposed:

\subsubsection{QKD with imperfect sources}

The source is typically less vulnerable to attacks because Alice can prepare her quantum states in a protected environment, and we expect that she can characterise her source. Therefore flaws in the state preparation can be incorporated into the security proof.

Loss-tolerant protocols has been proposed \cite{bib:PhysRevA.90.052314}, and further developed by \cite{bib:PhysRevA.92.032305}, where decoy state\index{Decoy states} QKD with tight finite-key security has been employed. A wide range of imperfections with the laser source has been taken into account in \cite{bib:mizutani2015finite}, including intensity fluctuation. A security proof has shown that perfect phase randomisation is also not necessary \cite{bib:cao2015discrete}.

This provides strong evidence that secure quantum communication with imperfect sources is feasible \cite{bib:diamanti2016practical}. Intuitively, QKD with imperfect sources are feasible, because by assuming that the states prepared by Alice are qubits, then Eve cannot unambiguously discriminate Alice's states \cite{bib:diamanti2016practical}.

\subsubsection{Measurement-device-independent QKD}

To prove security for Bob's measurement device is more problematic, since Eve has complete access to the quantum channel and she can send any signal.

One candidate for a long-term solution to side-channel attacks\index{Side-channel attacks} is device-independent (DI) QKD \cite{bib:PhysRevLett.98.230501}. It relies on the violation of a Bell inequality\index{Bell inequality} \cite{bib:hensen2015loophole}, and the security can be proven without knowledge of the implementation. However, the expected secure key-rate is low even at short distances. A more practical approach is measurement-device independence (MDI) \cite{bib:PhysRevLett.108.130503} QKD, which is immune to side-channel attacks against the measurement device. Here the device is treated as a black box, and does not need to be trusted. An important assumption for MDI QKD is that Eve cannot interfere with the state preparation process.

Another candidate such as detector-device-independent \cite{bib:lim2014detector, bib:PhysRevA.92.022337} QKD has been designed to take advantage of both the strong security of MDI-QKD, with the efficiency of conventional QKD. However, DDI-QKD has been shown to be vulnerable to attacks \cite{bib:PhysRevLett.117.250505}.

The MDI-QKD protocol has been extended to the continuous variable framework. However, this system requires homodyne detectors with efficiency $>85\%$, and a reliable phase reference between Alice and Bob.

We have discussed some significant remaining challenges in QKD. These range from theoretical security proofs to hardware developments. Advances in QKD will not only enable point-to-point quantum communication, but have implications for a range of network applications, such as quantum secret sharing \cite{bib:cleve1999share, bib:PhysRevA.61.042311, bib:PhysRevA.71.044301}, blind quantum computing \cite{bib:broadbent2009universal, bib:barz2012demonstration}, quantum anonymous broadcasting \cite{bib:christandl2005quantum}, and many more.

As remarked in \cite{bib:diamanti2016practical}, ``determining the exact power and limitations of quantum communication is the subject of intense research efforts worldwide. The formidable developments that can be expected in the next few years will mark important milestones towards the quantum internet of the future."