### 2018-10-09 Sections/statement_of_contributions.tex

%
% Statement of Contributions
%

\section*{Statement of contributions}

\begin{itemize}
\item Peter P. Rohde conceived, directed and edited this project, developed the original ideas, and prepared most of the manuscript.
\item Zixin Huang wrote Secs.~\ref{sec:telescopy}, \ref{sec:CV_QC}, \ref{sec:essay_future_QKD}.
\item Zuen Su \& Heliang Huang prepared Part~\ref{part:SotA}.
\item William J. Munro prepared Sec.~\ref{sec:rep_net}. 
%\item Ryan L. Mann contributed to Sec.~\ref{sec:protocols_quant_int}.
\item Scott Harrison wrote Secs.~\ref{sec:econ_prop}, \ref{sec:policy} \& \ref{sec:game_theory}.
\item Tim Byrnes \& Jonathan P. Dowling wrote Sec.~\ref{sec:quant_space_race}.
\item Nana Liu \& Chris Ferrie wrote Sec.~\ref{sec:quantum_mind}.
\item Nana Liu wrote Sec.~\ref{sec:verification}.
\item All authors played active roles in the discussions that developed this work. 
\end{itemize}

\comment{Add everyone else}
\comment{Put in same order as authorship}