### 2018-10-21 Sections/statement_of_contributions.tex

%
% Statement of Contributions
%

\section*{Statement of contributions}

\begin{itemize}
\item Peter P. Rohde conceived, directed and edited this project, developed the original ideas, and prepared most of the manuscript.
\item Zixin Huang wrote Secs.~\ref{sec:telescopy}, \ref{sec:CV_QC}, \ref{sec:essay_future_QKD} \& \ref{sec:NISQ}.
\item Heliang Huang \& Zuen Su prepared Part~\ref{part:SotA}.
\item Chris Ferrie \& Nana Liu prepared Sec.~\ref{sec:quantum_mind}.
\item Simon Devitt prepared Secs.~\ref{???}.
\item Rohit Ramakrishnan prepared Sec.~\ref{???}.
\item Si-Hui Tan \& Atul ??? contributed to Sec.~\ref{???}.
\item Nana Liu wrote Sec.~\ref{sec:verification}.
\item Scott Harrison wrote Secs.~\ref{sec:econ_prop}, \ref{sec:policy} \& \ref{sec:game_theory}.
\item Darryl Vietch contributed to Sec.~\ref{???}.
\item Chandra Radamankrishnan contributed to Sec.~\ref{???}.
\item Jonathan P. Dowling contributed to editing.
\item Tim Byrnes \& Jonathan P. Dowling wrote Sec.~\ref{sec:quant_space_race}.
\item William J. Munro prepared Sec.~\ref{sec:rep_net}. 
\item All authors played active roles in the discussions that developed this work. 
\end{itemize}