### 2018-11-07 Sections/the_quantum_sneakernet.tex
Sneaker net boat fig

%
% The Quantum Sneakernet
%

\section{The quantum Sneakernet\texttrademark}\index{Sneakernet}

\comment{To do. Simon Devitt.}

\begin{figure}[!htbp]
\if 3\pubmode
\includegraphics[width=0.7\textwidth]{sneakernet_boat}
\else
\includegraphics[width=0.475\textwidth]{sneakernet_boat}
\fi
\caption{\comment{To do}}\label{fig:sneakernet_boat}	
\end{figure}