### 2018-11-08 Sections/the_quantum_sneakernet.tex
Sneakernet fig

%
% The Quantum Sneakernet
%

\section{The quantum Sneakernet\texttrademark}\index{Sneakernet}

\comment{To do. Simon Devitt.}

\begin{figure}[!htbp]
\if 3\pubmode
\includegraphics[width=0.7\textwidth]{sneakernet_boat}
\else
\includegraphics[width=0.475\textwidth]{sneakernet_boat}
\fi
\caption{In the quantum Sneakernet\texttrademark entanglement distribution protocol, rather than using conventional flying qubits, we take highly error-corrected stationary qubits and physically transport them over long distances as freight. The error correction must be sufficient to maintain coherence over the time-scale of the journey, which could be anything from hours (when flying) to weeks (by sea\footnote{Or indefinitely if the destination is Australia.}). The mode of transport only transports one half of an error-corrected Bell-pair, transferring the initial entanglement between source and vessel, $\ket{\Psi^+}_{A,B}$, to be between source and destination, $\ket{\Psi^+}_{A,C}$. Because all Bell-pairs are identical and infinitely reproducible, latency presents no problems. For example, when using a Bell-pair to teleport a qubit over long distances, when the Bell-pair arrives is unimportant, as long as it's available at the time of teleportation. Thus, our Bell-pair cargo carriers can simply operate in the background, transporting Bell-pairs with as much bandwidth as possible, which may then be buffered by the recipient until required.}\label{fig:sneakernet_boat}	
\end{figure}