### 2018-11-15 Sections/statement_of_contributions.tex

%
% Statement of Contributions
%

\section*{Statement of contributions}

\begin{itemize}
	\item Peter P. Rohde conceived, directed and edited this project, developed the original ideas, and prepared most of the manuscript.
	\item Zixin Huang prepared Secs.~\ref{sec:telescopy}, \ref{sec:CV_QC}, \ref{sec:essay_future_QKD}, \ref{sec:attacks_QKD}, \ref{sec:shallow_circs} \& \ref{sec:NISQ}.
	\item Heliang Huang \& Zuen Su prepared Part.~\ref{part:SotA}.
	\item Chris Ferrie \& Nana Liu prepared Sec.~\ref{sec:quantum_mind}.
	\item Simon Devitt prepared Secs.~\ref{???}.
	\item Rohit Ramakrishnan prepared Sec.~\ref{???}.
	\item Si-Hui Tan \& Atul Mantri contributed to Sec.~\ref{???}.
	\item Nana Liu prepared Sec.~\ref{sec:verification}.
	\item Scott Harrison prepared Secs.~\ref{sec:econ_prop}, \ref{sec:policy} \& \ref{sec:game_theory}.
	\item Darryl Vietch contributed to Sec.~\ref{???}.
	\item Chandrashekar Radhakrishnan prepared Sec.~\ref{sec:artificial_atoms}.
	\item Jonathan P. Dowling contributed to editing.
	\item Tim Byrnes \& Jonathan P. Dowling prepared Sec.~\ref{sec:quant_space_race}.
	\item William J. Munro prepared Sec.~\ref{sec:rep_net}. 
	\item All authors played active roles in the discussions that developed this work. 
\end{itemize}