### 2018-11-29 Sections/about_the_authors.tex
Added Heliang section. Not edited

%
% About the Authors
%

\section*{About the authors}

%
% Peter Rohde
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_peter_rohde}
\end{center}

Peter Rohde is an ARC Future Fellow in the Centre for Quantum Software \& Information at the University of Technology Sydney, Australia (UTS:Q$|$SI$\rangle$), and associate member of the Hearne Institute for Theoretical Physics at Louisiana State University. He obtained a Bachelor of Computer Systems Engineering with First Class Honours, and a PhD in theoretical physics at the University of Queensland. He has worked at highly-acclaimed international institutes, including the University of Oxford, University of Queensland, Institute for Molecular Biosciences and Max-Planck Institute for the Science of Light, with over 60 publications and 1,500+ citations in quantum optics, quantum computing, quantum information theory, ecology, and politics. His theoretical proposals have inspired several world-leading experimental efforts, including time-bin encoded \textsc{BosonSampling} and sub-shot-noise limited quantum metrology. In his spare time he is a musician, composer, public speaker, libertarian political activist, charity worker, mountaineer and adrenaline junkie. He regularly speaks at prominent political and scientific outreach events, features in newspaper and magazine articles, and conducts radio interviews. He is a very naughty boy and was expelled from preschool\index{Montessori}.

%
% Heliang Huang
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_heliang_huang}
\end{center}

He-Liang Huang is a Postdoctoral Fellow at the University of Science \& Technology China (USTC). His research interests include secure cloud quantum computing, quantum big data analysis, and the physical implementation of quantum computing architectures, in particular using linear optical and superconducting systems. As an experimentalist he has been a part of several groundbreaking experimental efforts, most notably in the demonstration of world-leading multi-photon optical quantum information processing protocols.

%
% Zuen Su
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_zuen_su}
\end{center}

Zu-En Su is a PhD candidate at the University of Science \& Technology China.

\comment{Complete this section}

%
% Zixin Huang
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_zixin_huang}
\end{center}

Zixin Huang completed her PhD in physics at the University of Sydney, Australia. She is currently a postdoctoral fellow at the University of Sheffield, UK.

\comment{Complete this section}

%
% Chandra Radhakrishnan 
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_chandra_radhakrishnan}
\end{center}

Chandrashekar Radhakrishnan is a Postdoctoral Fellow and Bright Overseas Global Corporation Scholar at New York University, Shanghai, Chibna. He is a recipient of the National Natural Science Foundation of China, International Young Scientist research grant. Before Joining NYU Shanghai he worked as a Postdoctoral Fellow at the Institute of Mathematical Science, India and the National Chung Hsing University, Taiwan. He obtained his Master of Science in Physics from the Indian Institute of Technology Madras and a PhD in theoretical physics from University of Madras. He has worked on a variety of fields, including statistical mechanics, mathematical physics and quantum information theory. His fundamental interest in quantum information theory lies in investigating quantum resources alternative to entanglement, which can be used for quantum communication and quantum computation. He has active collaborations with experimentalists, through which he tries to verify his theoretical results. He also has some interest in the application of statistical mechanics to biological problems.

%
% Simon Devitt
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_simon_devitt}
\end{center}

\comment{Simon Devitt}

%
% Si-Hui Tan
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_sihui_tan}
\end{center}

Si-Hui Tan is a Research Scientist at the Singapore University of Technology \& Design. She obtained a Bachelor of Science with Honours from Caltech, and subsequently a PhD from MIT. She has received numerous awards, including the MIT Presidential Fellowship.

\comment{Complete this section}

%
% Nana Liu
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_nana_liu}
\end{center}

Nana Liu is a Postdoctoral Research Fellow at the Singapore University of Technology \& Design.

\comment{Complete this section}

%
% Rohit Ramakrishnan
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_rohit_ramakrishnan}
\end{center}

Rohit Ramakrishnan is a researcher at the Indian Institute of Science.

\comment{Complete this section}

%
% Jon Dowling
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_jon_dowling}
\end{center}

Jon Dowling is Professor and Hearne Chair of Theoretical Physics, and Co-Director of the Hearne Institute for Theoretical Physics at Louisiana State University. He has a highly-acclaimed research career, with over 450 publications and 15,000 citations. He is author of the popular book \textit{Schr\"odinger's Killer App: Race to Build the World's First Quantum Computer}.

\comment{Complete this section}

%
% Tim Byrnes
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_tim_byrnes}
\end{center}

Tim Byrnes....

\comment{Complete this section}

%
% Bill Munro
%

\begin{center}
\includegraphics[clip=true, width=0.475\textwidth]{photo_bill_munro}
\end{center}

Bill Munro obtained his PhD at the University of Waikato, and has since held numerous prestigious fellowships and positions. He is currently Senior Research Scientist and leader of the theoretical quantum physics research group at NTT Japan.

\comment{Complete this section}