### 2018-11-29 Sections/roadmap.tex
roadmap

%
% Roadmap - The way forward
%

\part{Roadmap -- The way forward}\index{Roadmap}

\dropcap{W}{e} have in this work presented a vast number of ideas, protocols, applications and implications for future quantum technology, mediated by the quantum internet. The relationship between all these concepts is complex, with a large number of interdependencies. In Fig.~\ref{fig:roadmap} we summarise this via a mind-map, capturing these relationships to show the logical flow of developments that will need to take place to reach various technological destinations.

\begin{figure*}
	\includegraphics[width=\textwidth]{roadmap}
	\caption{Relationships and dependencies in the development and deployment of the major concepts in the quantum internet.}\label{fig:roadmap}
\end{figure*}