### 2018-12-04 Sections/roadmap.tex

%
% Roadmap - The way forward
%

\part{Roadmap -- The way forward}\index{Roadmap}

\dropcap{W}{e} have in this work presented a vast number of ideas, protocols, applications and implications for future quantum technology, mediated by the quantum internet. The relationship between all these concepts is highly complex, with a large number of interdependencies. In Fig.~\ref{fig:roadmap} we express this as a mind-map timeline\index{Mind-map}\index{Timelines}, capturing these relationships to show the logical flow of developments that will need to take place to reach various technological milestones and destinations.

\begin{figure*}
	\includegraphics[clip=true, angle=-90, width=\textwidth]{roadmap}
	\caption{Relationships and dependencies in the development and deployment of the major concepts in the quantum internet, showing the logical flow of developments that will need to take place to reach various technological milestones and destinations.}\label{fig:roadmap}
\end{figure*}