### 2018-12-29 Sections/wine_pairings.tex
appendix

\section*{Recommended wine pairings}\index{Wine pairings}

\dropcap{L}{ike} any good meal, any good book chapter is best paired with a suitably-chosen wine that compliments its unique notes and characteristics. Here we suggest several contemporary, budget-friendly, Australian wines, hand-selected by the authors, that perfectly match the subtle undertones of each part in the book.

\begin{itemize}
\item Part.~\ref{part:introduction}: Howard Park Flint Rock Shiraz, 2015, Great Southern.
\item Part.~\ref{part:mathematical_foundations}: Helen \& Joey Inara Pinot Noir, 2016, Yarra Valley.
\item Part.~\ref{part:class_net}: Leconfield Merlot, 2016, Coonawarra.
\item Part.~\ref{part:quant_net}: Lou Miranda Leone Cabernet Sauvignon, 2015, Barossa Valley.
\item Part.~\ref{part:protocols}: Kaesler Stonehorse Grenache Shiraz, Mourvedre, 2014, Barossa.
\item Part.~\ref{part:ent_dist}: Ministry of Clouds Tasmania Chardonnay, 2016, Tamar Valley.
\item Part.~\ref{part:quant_crypto}: Piccadilly Valley Pinot Noir, 2016, Piccadilly Valley.
\item Part.~\ref{part:QC}: The Story R. Lane Vintners Westgate Vineyard Shiraz, 2015, Grampians.
\item Part.~\ref{part:cloud_QC}: d'Arenberg Artisans The Money Spider Roussanne, 2017, McLaren Vale.
\item Part.~\ref{part:economics}: Payten and Jones Major Kong Shiraz, 2015, Yarra Valley.
\item Part.~\ref{part:SotA}: Two Rivers Stones Throw Semillon, 2018, Hunter Valley.
\item Part.~\ref{part:future_beyond}: Penfolds Reserve Bin 17A Chardonnay, 2017, Adelaide Hills.
\item Part.~\ref{part:roadmap}: Watershed Senses Sauvignon Blanc, 2016, Margaret River.
\item Part.~\ref{part:essays}: Peter Lehmann Wines Wigan Riesling, 2015, Eden Valley.
\item Part.~\ref{part:the_end}: Giant Steps Yarra Valley Pinot Noir, 2017, Yarra Valley.
\end{itemize}

\textit{--- Don't drink and derive.}