### 2019-01-05 Sections/essay_the_quantum_mind.tex
+quotes, formatting

\section{The quantum mind}\index{Quantum mind}\label{sec:quantum_mind}

\sectionby{Nana Liu}

\famousquote{If computers get too powerful, we can organise them into committees. That'll do them in.}{Unknown}
\newline

\comment{To do: quantum machine learning and AI. QCs improve training process. Implications of strong AI for humanity.}

\comment{Selfishness arises naturally in evolution as a simple winning strategy in game theory. Need to make sure quantum AI doesn't become selfish. Must become collaborative/altruistic rather than selfish, so that it serves us rather than itself.}