### 2019-01-05 Sections/mathematical_foundations.tex

\famousquote{I come in peace. I didn't bring artillery. But I'm pleading with you, with tears in my eyes: If you fuck with me, I'll kill you all.}{General James ``Mad Dog'' Mattis}
\newline

\dropcap{T}{o} make the upcoming discourse accessible to those from a non-quantum background, we will briefly level the playing-field upfront by presenting an overview of the key concepts and mathematical formulations for quantum mechanics and quantum optics, both relied upon heavily throughout this work. All of this mathematical formulation fits within the confines of familiar basic undergraduate linear algebra\index{Linear algebra}\index{Quantum mechanics}\index{Quantum optics}.

This part ought provide the mathematical background necessary to follow the upcoming discussions. Those already familiar with the laws and formulation of quantum mechanics and quantum optics can calmly skip this part, without risk of offending the authors.

However, rather than just present the mathematical formulations and expect you to reverently accept the maths as the truth, the light, and the way, we will try as best we can to inject some intuition and meaningful interpretation of the mathematics into our tutorial, so that hopefully the reader is able to walk away with not just the ability to follow basic quantum mechanical derivations, but also grasp the intuition of the underlying physical reality.