### 2019-05-05 Sections/intro_to_QIT.tex
QIT + 1 added bib

%
% Introduction to quantum information theory
%
\clearpage

\section{Brief introduction to quantum information theory}\index{Quantum information theory}

\famousquote{When you find yourself in a room surrounded by your enemies you tell yourself, `I am not locked in here with you, you are locked in here with me'. This is the kind of mindset you should have if you want to succeed in life. Get rid of that victim mentality.}{Bruce Lee}

\subsection{Probability, information and classical correlation measures}

% The fundamental unit in Shannon information is the \textit{bit}.

``Information theory begins with the observation that there is a fundamental link between probabilities and information"\cite{barnett2009quantum}.

Suppose that there are two parties trying to communicate, conventionally named Alice and Bob. Alice 
sends information using an alphabet $\{x_j \}$, and each letter occurs with probability $p_j$. If
some letter occur less likely than others, then Bob will be more surprised receiving those, and one
might naturally expect that the information content of those events are higher.

This suggests a way to measure the \textit{information content} of an outcome, or how surprised Bob is:

\begin{align}
i(x_j) \equiv \log\left(\frac{1}{p(x_j)}  \right) =  -\log\left(p(x_j)  \right)
\end{align}

We can model Alice's source as a random variable $X$, whose outcomes
are $\{ x_j\}$ each occuring with probability $p_j$.
Then the expected information content of the source, or \textit{Shannon entropy} 
associated with the $X$ is defined as

\begin{align}\index{Shannon entropy}
H(X) = -\sum_i p_i\log_2(p_i).
\end{align}

Entropy plays a central role in information theory. The intuitive interpretation is that it quantifies an experimenter's uncertainty about $X$ before measuring it, and his expected information gain is $H(X)$ bits upon learning the outcome. The information $H(X)$ is zero if and only if one of the probabilities $p(x)$ is unity, with the others being zero. In this case the value of $X$ is already known and so there is no information to be gained from observing it.

 If there are two random variables $X$ and $Y$, the joint entropy of the two is simply given by,
\begin{align}\index{Joint Shannon entropy}
H(X,Y) =  -\sum_{x,y} p_{x,y}\log_2(p_{x,y}).
\end{align}

Note that the joint entropy $H(X,Y)= H(X)+H(Y)$ if and only if $X$ and $Y$ are independent,i.e. the occurence of one does not change the probability of the other.

\begin{figure}[hbt]
\includegraphics[clip=true, width=0.4\textwidth]{classical_mutual_info.pdf}
\captionspacefig \caption{\label{fig:mutual_info}}	
\end{figure}

If the two distributions are not independent, i.e. knowing something about $X$ reveals information on $U$, the two variables are said to be \textit{correlated}. Suppose Alice possesses the random variable $X$ and Bob has the random variable $Y$, the \textit{mutual information} specifies the number of bits in common between the two distributions. Equivalently, this represents the maximum number of bits that one party can learn about the other just by inspecting their own information.

For two classical distributions, the classical mutual information is given by,
\begin{align}\index{Mutual information}
I(X;Y) = H(X) + H(Y) - H(X,Y).
\label{classical_mutual_info}
\end{align}

It is a measure of the correlation between the events $X$ and $Y$. The quantity in Eq.~\eqref{classical_mutual_info} is important because it upper bounds the amount of information Alice and Bob can
reliably communicate. For a discrete random variable with $d$ possible outcomes, the maximum mutual information between $X$ and $Y$ is $I(X,Y)=\log_2 d$ bits.

Another quantity of interest is the \textit{conditional entropy}, which describes the entropy of $y$ should $X$ is known. The entropy of $Y$ conditioned on $X$ is given by
\begin{align}
H(X|Y) &= H(X,Y)- H(Y) \\
       &=- \sum_{x,y} p(x,y) \log_2 \frac{p(x,y)}{p(x)}.
\end{align}
\noindent Classically the conditional entropy is always positive, but as we will see later, this is not the case if quantum states are involved.

The relationship between the quantities mentioned above is graphically summarised in Figure~\ref{fig:mutual_info}.

% 
% 
% 
% 
%

\subsection{Quantum correlation measures}

The von Neuman entropy\index{von Neuman entropy} \cite{bib:bengtsson2017geometry} for quantum density operators, $S(\hat\rho)$, is defined analogously, replacing probabilities with density operator eigenvalues,
\begin{align}\index{von Neuman entropy}
S(\hat\rho) &= - \sum_x \lambda_x \log_2 (\lambda_x) \nonumber \\
&= -\mathrm{tr}(\hat\rho\,\log \,\hat\rho),
\end{align}
where $\{\lambda\}$ is the eigenvalue spectrum of $\hat\rho$. This modification is logically justified, as the eigenvalues can be interpreted directly as a purely classical probability distribution of orthogonal states when the density operator is transformed into a basis with no coherences between basis states (i.e a diagonal basis or spectral decomposition\index{Spectral decomposition}). In that case, the Shannon and von Neuman entropies essentially have identical physical interpretations.

There are a few properties of the Von Neumann entropy which are often useful \cite{}:
\begin{enumerate}
	\item For pure states $\hat \rho = \ket{\psi}\bra{\psi}, S(\hat\rho) =0$.
	\item The Von Neumann entropy is unchanged by a unitary change of basis, 
			\begin{align}
			S(U\hat \rho U^\dagger) = S(\hat \rho)
			\end{align}
	\item For a state living on a Hilbert space of dimension $d$, $S(\hat\rho) \leq \log_2 d$, the equality is achieved when the quantum state is maximally mixed.
	\item For a given bipartite system $AB$
			\begin{align}
			S(\hat \rho_{AB}) \leq S(\hat \rho_A) + S(\hat \rho_B).
			\end{align}
			Equality holds if $\rho_{AB}= \rho_A \otimes \rho_B$.
\end{enumerate}

Analogously, the quantum mutual information\index{Quantum mutual information} for bipartite state $\hat\rho_{A,B}$ is defined as,
\begin{align}
I(A;B)_{\hat\rho} = S(\hat\rho_A) + S(\hat\rho_B) - S(\hat\rho_{A,B}),
\end{align}
using the von Neuman entropy.

The mutual information between two quantum states is invariant under local unitary transformations,
\begin{align}
I(A;B)_{\hat\rho} = I(\hat{U}_A\hat\rho_A \hat{U}_A^\dag; \hat{U}_B\hat\rho_B \hat{U}_B^\dag),
\end{align}
since the eigenvalue spectrum of a density operator is invariant under unitary transformations. Therefore, the mutual information represents the maximum amount of information Bob can learn about Alice's state under \textit{any} local operations.

Now, in sharp contrast with the classical case, the conditional entropy for quantum states
\begin{align}
H(A| B)_{\hat \rho} = S(\hat \rho_{AB}) - S(\hat \rho_B)
\label{cond_quant_ent}
\end{align}
\noindent can be negative! If $\rho_{AB}$ is a maximally entangled state, then 
$S(\hat \rho_{AB}) =0$, and $H(A|B) =- S(\hat \rho_{AB}) <0$.
Intuitively, this can be interpreted as the fact that bipartite maximally entangled states can be more strongly correlated than classically possible, and that knowing the state of $A$ reveals an ``unnatural" amount of information about the state of $B$.

A quantum process cannot increase the mutual information between two parties. This yields the \textit{data processing inequality}\index{Data processing inequality} that, for a sequence of channels \mbox{$X\to Y\to Z$},
\begin{align}\index{Data processing inequality}\label{eq:data_proc_ineq}
I(X:Z)&\leq I(X:Y), \nonumber \\
I(X:Z)&\leq I(Y:Z),
\end{align}
with equality if and only if the channel not specified in the identity on the right hand side (\mbox{$Y\to Z$} or \mbox{$X\to Y$} respectively) is unitary, i.e one of the links in the chain perfectly preserves information content. The progression is shown in Fig.~\ref{fig:data_proc_ineq}.

\begin{figure}[!htbp]
\includegraphics[clip=true, width=0.3\textwidth]{data_proc_ineq}
\captionspacefig \caption{\label{fig:data_proc_ineq}A sequence of events \mbox{$X\to Y\to Z$}. The data processing inequality states that the mutual information from beginning to end is upper-bounded by the mutual information between neighbouring stages, as per Eq.~(\ref{eq:data_proc_ineq}).}	
\end{figure}

The mutual information is defined as being between a particular known pair of states. Of course, in a quantum channel we have the flexibility to manipulate the input state and the measurement at the output. From this, we can then define the \textit{classical information of the channel}\index{Classical information of the channel}, $I_c(\hat\rho_{A,B})$ -- the maximum of the mutual information, optimised over \textit{all} possible measurement settings. Suppose $\hat\rho_{A,B}$ is shared between Alice and Bob, from which they would like to extract maximal correlations (i.e joint information). They measure using local measurement bases $\{\Lambda_A^x\}$ and $\{\Lambda_B^x\}$, yielding random variables $A$ and $B$. The quantum mutual information of the channel $\mathcal{E}$ is defined as,
\begin{align}
I_c(\mathcal{E}) = \max_{\rho,\{\Lambda_A^x\},\{\Lambda_B^x\}} I(A;B)_\rho. 
\label{eq:channel_quant_mutual}
\end{align}

Eq.~\eqref{eq:channel_quant_mutual} quantifies the maximum amount of correlation that can be established between input and output states that can be achieved across the channel. This dictates the physical upper bound on the achievable bitrate (classical communication) or bandwidth of the channel.

\vspace{5mm}
% \paragraph{Quantum correlation measures}
All of the correlations discussed above concern classical information.
We now move on to discuss quantum correlations. Whilst classical correlation measures have a clear operational meaning, quantum channel capacities are less so -- these quantities can be interpreted as how well the channel preserves the quantum state.

The conditional entropy for quantum states can be negative is significant, in fact, it has its own name and notation, the \textit{coherent information}
\begin{align}
I(A\rangle B) = H(B)_{\hat\rho} - H(A,B)_{\hat\rho},
\end{align}

\noindent For a maximally entangled state, the coherent information is equal to one. It measures to what extent we know less about a subsystem compared to the whole.