### 2019-06-26 Sections/verification_of_quantum_computing.tex

%
% Verification Of Cloud Quantum Computing
%

\section{Verifiable quantum computing} \index{Verification}

\sectionby{(in progress..)}

Consider a scenario where....

%{Physics and computer science view}

\subsection{Benchmarking}
References???? 
\subsection{Formal Definitions}

\begin{definition}[Stand-alone Verification]
\label{def:vbqc}
Let $\mathcal{P}$ be a delegated quantum computing protocol between a client and a server.  Let the client's state be $|\psi'\rangle =  |\psi\rangle \otimes |\text{flag}\rangle$, where $|\psi\rangle$ is the input state to the protocol and $|\text{flag}\rangle$ denotes whether the client accepts ($|\text{flag}\rangle = |\text{acc}\rangle$) or rejects ($|\text{flag}\rangle = |\text{rej}\rangle$) at the end of the protocol. Similarly, consider the quantum channel $\mathsf{Enc}_s$ (short form for encoding circuit) acting on the client's quantum state depending on the private random string $s$, sampled by the client from some distribution $p(s)$.  We also denote $P_\text{honest}$ as the CPTP map acting on the client's state when the server follows the steps of the protocol. The incorrect subspace is written as: 
\begin{equation}
P^s_\text{inc} = (I - |\psi_o^s\rangle \langle \psi_o^s|) \otimes |\text{acc}^s\rangle \langle \text{acc}^s|
\end{equation}
as a projection onto the orthogonal complement of the correct output: 
\begin{equation}
|\psi_o^s\rangle \langle \psi_o^s | = {\text{Tr}}_{\text{flag}}(P_\text{honest}(\mathsf{Enc}_s(|\psi'\rangle \langle \psi'|)))
\end{equation}
and on the acceptance for the \text{flag} state: 
\begin{equation}
|\text{acc}^s\rangle \langle \text{acc}^s| = {\text{Tr}}_{|\psi\rangle}(P_\text{honest}(\mathsf{Enc}_s(|\psi'\rangle \langle \psi'|)))
\end{equation}
We say that such a protocol is \textit{$\epsilon$-verifiable} (with $0\leq \epsilon \leq 1$), if for any action $P$ of the the server we have
\begin{equation} \label{eq:verif}
\text{Tr} \big( \sum_s p(s) P^s_\text{inc} P(\mathsf{Enc}_s(|\psi'\rangle)) \big) \leq \epsilon
\end{equation}
\end{definition}

Roughly speaking, Eq~\eqref{eq:verif} states that the probability for the client accepting the incorrect output (given by a malicious server) must be bounded by $\epsilon$. In other words, the client would be able to detect almost all the times whenever the server deviates from the protocol. We now briefly summarize the protocols with the settings where client and server's resources vary.

\subsubsection{Composable definition}

\subsection{Parameters}
\subsubsection{Completeness}
\subsubsection{Soundness}
\subsubsection{Interaction}
\subsubsection{Complexity of Verifiers and Prover}
AG'17, ...

\subsection{Models}
\subsubsection{Circuit Model}
Broadbent
\subsubsection{MBQC}
FK
\subsubsection{Hamiltonian Model}
Joe, Tomoyuki,,,history state

\subsection{Techniques}
There are several techniques ...mainly ..
\subsubsection{Trap based}
 FK, ...
\subsubsection{Stabilizer}
Tomoyuki ...
\subsubsection{Self testing}
Mckague, ...
\subsubsection{Certification protocols}
scheme by Eisert ??? Reliable quantum certification for photonic quantum technologies, (Hamiltonian Learning and Certification Using Quantum Resources https://arxiv.org/pdf/1309.0876.pdf ), weak membership certification...
\subsubsection{Post-hoc}
Joe...
\subsubsection{Commit and Reveal}
mahadev, CCKW, GV ...
\subsubsection{Authentication}
polynomial codes, trap codes, ...QPIP Aharonov
\subsubsection{Recursive Sampling, Sampling...}
scheme of tomoyuki, joe tommaso,

\subsubsection{Other schemes}
Arrighi and Salvail weakly random function? randomness certification, Tomoyuki Rational proofs

\subsection{Non-universal quantum computing schemes }
IQP, Boson sampling, Tomoyuki..
\subsection{Classical verification protocols..}

\subsubsection{complexity theory} 
Interactive proofs (IP = PSPACE)...

\subsection{Schemes}
prepare and send, receive and measure, classical client - commit and reveal....

\subsection{Example}
FK or post-hoc protocol ...