### 2019-08-13 SECTION - Error teleportation & swapping/error_teleportation_swapping.tex
EUROPE RETURN STAMP

\documentclass[aps,pra,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}

\begin{document}

\bibliographystyle{apsrev}

%
% Title
%

\title{}

%
% Authors
%

\author{Peter P. Rohde}

\frenchspacing

\maketitle

\section{Error propagation via entanglement}

\subsection{Entanglement-symmetric errors}

\begin{align}
	\ket{\Phi^+}_{A,B}=\frac{1}{\sqrt{2}}(\ket{0}_A\ket{0}_B + \ket{1}_A\ket{1}_B),
\end{align}

\begin{align}
	\hat{Z}_A \ket{\Phi^+}_{A,B} &= \hat{Z}_B \ket{\Phi^+}_{A,B} = \nonumber\\
	\ket{\Phi^-}_{A,B} &= \frac{1}{\sqrt{2}}(\ket{0}_A\ket{0}_B - \ket{1}_A\ket{1}_B),
\end{align}

\begin{align}
	\mathcal{E}^A_p(\ket{\Phi^+}_{A,B}\bra{\Phi^+}_{A,B}) = \mathcal{E}^B_p(\ket{\Phi^+}_{A,B}\bra{\Phi^+}_{A,B}),
\end{align}
where $\mathcal{E}$ is either a single-qubit dephasing, bit-flip or depolarising channel. Note that this type of symmetry does not apply in general. For example, it is intuitively obvious upon inspection that such symmetries do not apply for amplitude damping or loss channels, for example.

\textit{entanglement-symmetric errors}

\begin{itemize}
\item Dephasing channels.
\item Bit-flip channels.
\item Depolarisation channels.	
\end{itemize}

\begin{figure}
\caption{To do.}\label{fig:error_symmetry}	
\end{figure}

When both qubits within the Bell pair are subject to independent entanglement-symmetric errors, we can simplify the description of the noisy state by shifting one side's errors over to the other, and appropriately combining them into an appropriately reparameterised single-qubit error channel acting exclusively only one qubit (the other treated as ideal, unaffected by any channel). That is, a Bell pair where both qubits are independently subject to distinct entanglement-symmetric error channels, can be fully specified using just a single, combined error rate, acting arbitrarily on one qubit or the other (whichever is mathematically convenient).

For Bell pairs, and any entanglement-symmetric error channels $\mathcal{E}$, the following representations for the two-qubit error channels are mathematically equivalent,
\begin{align}
	\mathcal{E}^A_{p_A}\otimes \mathcal{E}^B_{p_B} &\longleftrightarrow \mathcal{E}^A_{p_B}\otimes \mathcal{E}^B_{p_A} \nonumber\\
	&\longleftrightarrow (\mathcal{E}^A_{p_A}\circ \mathcal{E}^A_{p_B})\otimes \hat{I}^B \nonumber\\
&\longleftrightarrow \hat{I}^A\otimes (\mathcal{E}^B_{p_A}\circ \mathcal{E}^B_{p_B}).
\end{align}

Same symmetry in reversed time direction for Bell measurements. That is, a Bell measurement between two unentangled qubits allows entanglement-symmetric errors to hop across the divide from one side to the other arbitrarily.

\subsection{Error teleportation}

\begin{figure}
\caption{To do.}\label{fig:error_teleport}	
\end{figure}

\subsection{Entanglement swapping}

Applying this identity recursively yields an extremely powerful process for simplification. Specifically, for a linear chain of alternating Bell pair distributors and entanglement swappers (Fig.~\ref{fig:error_swapping}), as might be used to concatenate a sequence of short entanglement links into a unified long-range link, this implies that entanglement-symmetric error channels may be accommodated for in an extremely trivial manner -- simply accummulate them all together into a composed single-qubit operation, which acts on one of the qubits of all that remains, a single long-range Bell pair. Specifically, for $n$ Bell pairs, each subject two independent single-qubit error channels, interspersed with entanglement swappers (Bell measurements), the residual Bell pair is characterised entirely by the cumulative single-qubit channel,
\begin{align}
\mathcal{E}_{p_\mathrm{joint}} = \prod_{i=1}^{2n} \mathcal{E}_{p_i}.
\end{align}

In the special case of depolarising channels, the accumulation of the parameters is simply multiplicative in nature,
\begin{align}
p_\mathrm{joint} = \prod_{i=1}^{2n} p_i.	
\end{align}

\begin{figure}
\caption{To do. (a) and (b) for 1 and $n$ swappers}\label{fig:error_swapping}
\end{figure}

\section{Cost vector analysis}

\subsection{Cost teleportation}

\bibliography{paper}

\section{Entanglement purification}

\begin{align}
F' = \frac{F_A F_B}{F_A F_B + (1-F_A)(1-F_B)}
\end{align}

(double check) Note that provided both inputs exceed the \mbox{$F=1/2$}, the protocol guarantees that the fidelity of the purified pair exceed that of \textit{both} inputs -- things necessary improve over what we began with.
\begin{align}
F' \geq F_{A,B} \geq 1/2.
\end{align}

\end{document}