### 2020-09-01 cambridge.tex
almost done editing.

%\NeedsTeXFormat{LaTeX2e}[1996/06/01]

\documentclass[multi,spanningrule]{cambridge6A}

\def\authorpage{0}
\def\abstractpage{0}
\def\dedicationpage{0}
\def\sketchesmode{0} 	
\def\seriousmode{1} 			
\def\maincontribstyle{0}
\def\QIPlink{0}
\def\parainTOC{0}
\def\includelicense{0}
\def\doublecol{0}

\input{Sections/packages}
\input{Sections/commands_and_definitions}

  \usepackage{natbib}

  \usepackage{rotating}
  \usepackage{floatpag}
  \rotfloatpagestyle{empty}

 %\usepackage{amsmath}
 %  \usepackage{amsthm}
 % \usepackage{graphicx}

% for a single index
% \usepackage{makeidx}
 \makeindex

 \newcommand\cambridge{cambridge6A}

  \setcounter{tocdepth}{2}% the toc normally lists sections;
% for the purposes of this document, this has been extended to subsections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \includeonly{chap1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{document}

  \title[]{The Quantum Internet}

  \author{Peter P. Rohde}

  \frontmatter
  \maketitle
  \tableofcontents
 % \listoffigures
 % \listoftables
  \listofcontributors

  \mainmatter

%
% Introduction
%

\part{Introduction}\label{part:introduction}\index{Introduction}

\famousquote{Any sufficiently advanced technology is indistinguishable from magic.}{Arthur C. Clarke}
\newline

\famousquote{There may be babblers, wholly ignorant of mathematics, who dare to condemn my hypothesis, upon the authority of some part of the Bible twisted to suit their purpose. I value them not, and scorn their unfounded judgement.}{Nicolaus Copernicus}\index{Grandad}\index{Mansplaining}
\newline

\famousquote{Imagination is more important than knowledge.}{Albert Einstein}
\newline

\input{Sections/foreword}

\latinquote{Liber magnus est.}

\input{Sections/introduction}

\latinquote{Vincit qui se vincit.}

%\input{Sections/quantum_internet_project}

\sketch{sketch_2}

%
% Mathematical foundations
%

%\part{Mathematical foundations}\label{part:mathematical_foundations}\index{Mathematical foundations}

%\input{Sections/mathematical_foundations}
%\input{Sections/intro_to_QM}
%\input{Sections/intro_to_QO}
%\input{Sections/intro_to_QIT}

\latinquote{Alis volat prorilis.}

\sketch{sketch_3}

%
% Classical networks
%

\part{Classical networks}\label{part:class_net}\index{Classical networks}

\input{Sections/classical_networking_protocols}

\latinquote{Te futueo et caballum tuum.}

\input{Sections/mathematical_representation_of_quantum_networks}

\latinquote{Nil sine magno labore.}

\input{Sections/network_topologies}

\latinquote{Vitam regit fortuna, non sapientia.}

\input{Sections/network_algorithms}

\sketch{sketch_4}

%
% Quantum networks
%

\part{Quantum networks}\label{part:quant_net}\index{Quantum networks}

\input{Sections/quantum_networks}

\latinquote{Gladiator in arena consilium capit.}

\input{Sections/quantum_channels}

\latinquote{Bulla crustulum.}

\input{Sections/optical_encoding_of_quantum_information}

\latinquote{Audaces fortuna juvat.}

\input{Sections/errors_in_quantum_networks}

\latinquote{Igne natura renovatur integra.}

\input{Sections/costs_and_attributes_in_quantum_networks}

\latinquote{Flectere si nequeo superos, acheronta movebo.}

\input{Sections/routing_strategies}

\latinquote{Disce quasi semper victurus vive quasi cras moriturus.}

\input{Sections/interconnecting_and_interfacing_quantum_networks}

\latinquote{Vir sapit qui pauca loquitur.}

\input{Sections/optical_routers}

\latinquote{Semper inops quicumque cupit.}

\input{Sections/optical_stability_in_quantum_networks}

\latinquote{Dum inter homines sumus, colamus humanitatem.}

\sketch{sketch_5}

%
% Protocols for the quantum internet
%

\part{Protocols for the quantum internet}\label{part:protocols}\index{Protocols}

\input{Sections/protocols_for_the_quantum_internet}

\latinquote{Etiam capillus unus habet umbram.}

\sketch{sketch_6}

%
% Entanglement distribution
%

\part{Entanglement distribution}\label{part:ent_dist}\index{Entanglement!Distribution}

\input{Sections/entanglement_the_ultimate_quantum_resource}

\latinquote{Credo in unum deum.}

\input{Sections/quantum_repeater_networks}

\latinquote{Acta non verba.}

\input{Sections/the_irrelevance_of_latency}

\latinquote{Ceterum censeo carthaginem delendam esse.}

\input{Sections/the_quantum_sneakernet}

\latinquote{Veritas vos liberabit.}

\sketch{sketch_7}

%
% Quantum cryptography
%

\part{Quantum cryptography}\label{part:quant_crypto}\index{Quantum cryptography}

\input{Sections/cryptowars}

\latinquote{A mari usque ad mare.}

\sketch{sketch_8}

%
% Quantum computing
%

\part{Quantum computing}\label{part:QC}\index{Quantum computing}

\input{Sections/models_for_quantum_computation}

\latinquote{Aqua vitae.}

\input{Sections/quantum_algorithms}

\latinquote{Sic semper tyrannis.}

\input{Sections/physical_architectures_for_quantum_computing}

%\latinquote{Acta deos numquam mortalia fallunt.}

%\input{Sections/quality_of_service}

\latinquote{Ex luna scientia.}

\sketch{sketch_9}

%
% Cloud quantum computing
%

\part{Cloud quantum computing}\label{part:cloud_QC}\index{Cloud quantum computing}

\input{Sections/cloud_quantum_computing}

\latinquote{Timendi causa est nescire.}

\input{Sections/encrypted_cloud_quantum_computing}

\latinquote{Audentes fortuna iuvat.}

\input{Sections/verification_of_cloud_quantum_computing}

\latinquote{Semper fidelis.}

\sketch{sketch_10}

%
% Economics & Politics
%

\part{Economics \& politics}\label{part:economics}\index{Economics}\index{Politics}

\input{Sections/economics_of_the_quantum_internet}

\sketch{sketch_11}

\clearpage
%
%%
%% To the future & beyond
%%
%
%\part{To the future \& beyond}\label{part:future_beyond}
%
%\input{Sections/to_the_future_and_beyond}
%
%\latinquote{Excelsior.}
%
%\input{Sections/the_quantum_space_race}
%
%\latinquote{Volens et potens.}
%
%\input{Sections/quantum_genetic_medicine}
%
%\latinquote{Annus mirabilis.}
%
%\input{Sections/the_quantum_mind}
%
%\latinquote{Grandescunt aucta labore.}
%
%%\input{Sections/quantum_biology}
%
%\input{Sections/optimising_the_world}
%
%\latinquote{Ceteris paribus.}
%
%\sketch{sketch_13}
%
%%
%% Roadmap - The way forward
%%
%
%\input{Sections/roadmap}
%
%\latinquote{Aut inveniam viam aut faciam.}
%
%\sketch{sketch_14}
%
%
% Essays - A New Frontier
%

\part{Essays}\label{part:essays}\label{part:essays}\index{Essays}\index{A new frontier}

\input{Sections/a_new_frontier_towards_the_singularity}

\sketch{sketch_15}

%
% The end
%

\part{The end}\label{part:the_end}

\input{Sections/conclusion_the_vision_of_the_quantum_internet}

%
%\input{Sections/how_to_learn_more}
%
%\if 0\seriousmode
%	\input{Sections/epilogue}
%\fi
%
%\latinquote{Opus dei.}
%
%\input{Sections/acknowledgements}
%
%\latinquote{Dum spiro, spero.}
%
%\input{Sections/statement_of_contributions}
%
%\input{Sections/list_of_abbreviations}
%

\bibliography{quantum_internet, nana_ML}
\bibliographystyle{cambridgeauthordate}

%
%\input{Sections/index}
%
%%\input{Sections/about_the_authors}
%
%\sketch{sketch_final}
%
%\famousquote{We all die. The goal isn't to live forever, the goal is to create something that will.}{Chuck Palahniuk}
%

  \backmatter

  \addtocontents{toc}{\vspace{\baselineskip}}
  \theendnotes

  \cleardoublepage

\printindex

\end{document}